---
title: "EDA_LA_2"
author: "Reshma Itagi"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
githublink: https://github.com/ReshmaCode/EDA_LA2
dataset: Test_scores.csv
---
```{r}
df = read.csv('test_scores.csv')
head(df)
```

```{r}
library(dplyr)
library(ggplot2)
```

```{r}
temp = df %>% filter(df$school == "ANKYI")
temp.aov = na.omit(aov(temp$pretest ~ temp$posttest, data = temp))
temp.aov
```


```{r}
ggplot(df,aes(x=pretest, fill=n_student, y=posttest))+geom_col(position ="dodge")
```

# Creating a Scatter Plot
```{r}
plot(df$pretest, df$posttest)
# Scatter plot with base graphics
```

# Scatter plot with ggplot 2
```{r}
library(ggplot2)
ggplot(df, aes(x = n_student, y = pretest)) +geom_point()

ggplot(data = NULL, aes(x =df$pretest, y=df$posttest )) +
geom_point()
```

# Creating a Line Graph
```{r}
plot(df$pretest,type = "l")
```

# Line graph with base graphics
```{r}
plot(head(df$pretest),head(df$posttest),  type = "l")
points(head(df$pretest),head(df$posttest))
```
# Line graph with ggplot() and With points added to ggplot()
```{r}
ggplot(head(df,n=20), aes(x = pretest, y = posttest)) +geom_line()
ggplot(head(df,n=20), aes(x = pretest, y = posttest)) +geom_line()+geom_point()
```
# Creating a Bar Graph
```{r}
barplot(head(df$n_student), names.arg = head(df$pretest))
```
# Generate a table of counts
```{r}
barplot(table(df$pretest))
```
# Laoding ggplot2() package
```{r}
library(ggplot2)
```
# Bar graph of values. This uses the dataset data frame, with the
# "pretest" column for x values and the "posttest" column for y values.
```{r}
ggplot(df, aes(x = pretest, y = posttest)) +geom_col()
```
# Convert the x variable to a factor, so that it is treated as discrete
```{r}
ggplot(df, aes(x = factor(pretest), y = posttest)) +geom_col()
```

# Bar graph of counts. This uses the dataset data frame, with the "pretest" column for
# x position. The y position is calculated by counting the number of rows for
# each value of pretest.
```{r}
ggplot(df, aes(x = pretest)) +
geom_bar()
# Bar graph of counts
ggplot(df, aes(x = factor(pretest))) +
geom_bar()
```
# Creating a Histogram
```{r}
hist(df$pretest)
# Specify approximate number of bins with breaks
hist(df$pretest, breaks = 10)
```
#ggplot2 histogram with default bin width (left); With wider bins (right)
```{r}
ggplot(df, aes(x = pretest)) +
geom_histogram()
ggplot(df, aes(x = pretest)) +
geom_histogram(binwidth = 4)
```
# Creating a Box Plot
```{r}
boxplot(head(df$n_student,n=500), head(df$pretest,n=500))
```
# Make box plots for multiple variables, by combining the variables with interaction()
```{r}
ggplot(head(df,n=100), aes(x = interaction(pretest, posttest), y = n_student)) +
geom_boxplot()
```
# Making a Basic Histogram
```{r}
ggplot(df, aes(x = pretest)) +geom_histogram()
```
```{r}
ggplot(df, aes(x= posttest)) +
geom_histogram(binwidth = 5, fill = "white", colour = "black")
```
# Map pretest to fill, make the bars NOT stacked, and make them semitransparent
```{r}
ggplot(df, aes(x = pretest, fill ='pretest',colour='red' )) +
geom_histogram(position = "identity", alpha = 0.4)
```
# Making a Density Curve
```{r}
ggplot(df, aes(x = n_student)) +geom_density()
# expand_limits() increases the y range to include the value 0
ggplot(df, aes(x = pretest)) +geom_line(stat = "density") +expand_limits(y = 0)
```
# Density curve with a smaller and larger value of adjust:
```{r}
ggplot(df, aes(x = posttest)) +geom_line(stat = "density") +
geom_line(stat = "density", adjust = .25, colour = "red") +
geom_line(stat = "density", adjust = 2, colour = "blue")
```
# This draws a blue polygon with geom_density(), then adds a line on top
# Density curve with wider x limits and a semitransparent fill (left); In two
# parts, with geom_density() and geom_line() (right)
```{r}
ggplot(df, aes(x = pretest)) +
geom_density(fill = "blue", alpha = .2) +
xlim(35, 105)
ggplot(df, aes(x = pretest)) +
geom_density(fill = "blue", alpha = .2, colour = NA) +
xlim(35, 105) +
geom_line(stat = "density")
```
# Density curve overlaid on a histogram
```{r}
ggplot(df, aes(x = pretest, y = ..density..)) +geom_histogram(fill = "cornsilk", colour = "grey60", size = .2) +geom_density() +xlim(35, 105)
```
# Making Multiple Density Curves from Grouped Data
```{r}
data5 <- df %>%
mutate(n_student = as.factor(n_student)) # Convert n_student to a factor
# Map n_student to colour
ggplot(data5, aes(x = pretest, colour = "red")) +geom_density()
# Map n_student to fill and make the fill semitransparent by setting alpha
ggplot(data5, aes(x = pretest, fill = pretest)) +geom_density(alpha = .3)
```
# Making a Frequency Polygon
```{r}
ggplot(df, aes(x=pretest)) +
geom_freqpoly()
ggplot(faithful, aes(x = waiting)) +      
geom_freqpoly(binwidth = 4)           #controlling bin width
```
# Divide the x-axis range into 15 bins
```{r}
binsize <- diff(range(df$pretest))/15
ggplot(df, aes(x = pretest)) +
geom_freqpoly(binwidth = binsize)
```
# Making a Basic Box Plot
```{r}
ggplot(df, aes(x = factor(n_student), y = pretest)) +
geom_boxplot()
```
# Box plot with narrower boxes (1)
# With smaller, hollow outlier points(2)
```{r}
ggplot(df, aes(x = factor(n_student), y = pretest)) +geom_boxplot(width = .5)
ggplot(df, aes(x = factor(n_student), y = pretest)) +geom_boxplot(outlier.size = 1.5, outlier.shape = 21)
```

```{r}
ggplot(df, aes(x = 1, y = pretest)) +geom_boxplot() +scale_x_continuous(breaks = NULL) +theme(axis.title.x = element_blank())
```

# Adding notches to a box plot to assess whether the medians are different.
```{r}
ggplot(df, aes(x = factor(n_student), y = pretest)) +geom_boxplot(notch = TRUE)
```
# Adding means to box plot
```{r}
ggplot(df, aes(x = factor(n_student), y = pretest)) +geom_boxplot() +stat_summary(fun = "mean", geom = "point", shape = 23, size = 3,
fill = "white")
```
# Making a Violin Plot
```{r}
data6 <- ggplot(df, aes(x = pretest, y = n_student))
data6+geom_violin()
```
# A violin plot with box plot overlaid on it
```{r}
data6+geom_violin() +geom_boxplot(width = .1, fill = "black", outlier.colour = NA) +
stat_summary(fun= median, geom = "point", fill = "white", shape = 21,
size = 2.5)
```
# A violin plot with tails
```{r}
data6+geom_violin(trim = FALSE)
```
# Violin plot with area proportional to number of observations
```{r}
data6 +geom_violin(scale = "count")
```
# Violin plot with
```{r}
# More smoothing
data6+geom_violin(adjust = 2)
# Less smoothing
data6 +geom_violin(adjust = .5)

```

# Making a Dot Plot
```{r}
# Save a modified data set that only includes of males data for marks that contains > 50
data7 <- df %>%
filter(gender == "Male" & pretest > 50)
```
```{r}
# Create a base ggplot object using `data7`, called `data7_p` (for data7 plot)
data8 <- ggplot(data7,aes(x=pretest))
data8+geom_dotplot()
```
# Dot plot with stackdir = “center” (1)
# With stackdir = “centerwhole”(2)
```{r}
data8 +
geom_dotplot(binwidth = .25, stackdir = "center") +
scale_y_continuous(breaks = NULL) +
theme(axis.title.y = element_blank())
data8 +
geom_dotplot(binwidth = .25, stackdir = "centerwhole") +
scale_y_continuous(breaks = NULL) +
theme(axis.title.y = element_blank())
```
# Making Multiple Dot Plots for Grouped Data
```{r}
ggplot(df, aes(x = gender, y = n_student)) +
geom_dotplot(binaxis = "y", binwidth = .5, stackdir = "center")
```
# Dot plot overlaid on box plot
```{r}
ggplot(df, aes(x = gender, y = n_student)) +
geom_boxplot(outlier.colour = NA, width = .4) +
geom_dotplot(binaxis = "y", binwidth = .5, stackdir = "center", fill = NA)

```
# Dot plot next to box plot
```{r}
ggplot(df, aes(x = pretest, y = n_student)) +
geom_boxplot(aes(x = as.numeric(pretest) + .2, group = pretest), width = .25) +geom_dotplot(aes(x = as.numeric(pretest) - .2, group = pretest),binaxis = "y",binwidth = .5,stackdir = "center") +scale_x_continuous(breaks = 1:nlevels(df$pretest),labels=levels(df$pretest)
)
```
# Making a Density Plot of Two-Dimensional Data
```{r}
# Save a base plot object
 ggplot(df, aes(x = pretest, y = posttest))+geom_point() +stat_density2d()
# Contour lines, with "height" mapped to color
ggplot(df, aes(x = pretest, y = posttest))+geom_point() +stat_density2d()+stat_density2d(aes(colour = ..level..))

```
# With ..density.. mapped to fill (1)
# With points, and ..density.. mapped to alpha(2)
```{r}
# Map density estimate to fill color
ggplot(df, aes(x = pretest, y = posttest))+geom_point() +stat_density2d()+stat_density2d(aes(fill = ..density..), geom = "raster", contour = FALSE)
# With points, and map density estimate to alpha
ggplot(df, aes(x = pretest, y = posttest))+geom_point() +stat_density2d()+geom_point() +stat_density2d(aes(alpha = ..density..), geom = "tile", contour = FALSE)
```
# Plotting a Function
# # The data frame is only used for setting the range
```{r}
# The normal distribution
ggplot(data.frame(x = c(-3, 3)), aes(x = x))  + stat_function(fun = dnorm)
```
```{r}
# The t-distribution with df=2
ggplot(data.frame(x = c(-3, 3)), aes(x = x)) + stat_function(fun = dt, args = list(df = 2))

```










